---
title: "testing options of swarm"
output: html_notebook
---
Here we created small bash tests to test two options of swarm, the new network option and the fastidious option. 
You see a description of each test and what we expect, then the test itself and the output afterwards.

1. -network option:

option dumps the network of single-difference amplicons to a specified 
file in a sorted and tab-separated format (seed amplicon, neighbour 
amplicon). Use -n to get the full network, otherwise only links between 
amplicons where the abundance of the seed is higher or equal to the 
abundance of the neighbour is included.

```{r Tests for -j option}
#!/bin/bash

# test sequences with one difference (different abundances): we expect one link
printf ">s1_3\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 1 ? 0 : 1}' && echo success || echo failure

s1_3    s2_1
success

# test sequences with two differences (different abundances): we expect no link
printf ">s1_3\nA\n>s2_1\nCC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 0 ? 0 : 1}' && echo success || echo failure

success

# test sequences with one difference (same abundances): we expect two links
printf ">s1_1\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 2 ? 0 : 1}' && echo success || echo failure

s1_1    s2_1
s2_1    s1_1
success

# test sequences with two differences (same abundances):  we expect no link
printf ">s1_1\nA\n>s2_1\nCC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 0 ? 0 : 1}' && echo success || echo failure

success

# tab-separated format
printf ">s1_3\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
grep -q $'\t' && echo success || echo failure

success

# sorted by decreasing abundances within line (most abundant amplicon first)
printf ">s1_1\nA\n>s2_3\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit $1 == "s2_3" ? 0 : 1}' && echo success || echo failure

s2_3    s1_1
success

# sorted by decreasing abundances between lines (most abundant amplicon first)
printf ">s1_2\nA\n>s2_1\nC\n>s3_1\nTT\n>s4_3\nGT\n" | swarm -l /dev/null -o /dev/null -j - | \
cut -f 1 | tr "\n" " " | awk -F "[_ ]" '{exit $2 > $4 ? 0 : 1}' && echo success || echo failure

s4_3    s3_1
s1_2    s2_1
success

# binds from high to low abundances: we expect only one link if sequences have different abundance values
printf ">s1_3\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 1 ? 0 : 1}' && echo success || echo failure

s1_3    s2_1
success

# binds between equals: we expect double link if sequences have different abundance values
printf ">s1_1\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | tr "\n" " " | \
awk '{exit $1 == $4 && $2 == $3 ? 0 : 1}' && echo success || echo failure

s1_1    s2_1
s2_1    s1_1
success

# double links are from A to B and from B to A
printf ">s1_1\nA\n>s2_1\nC\n" | swarm -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 2 ? 0 : 1}' && echo success || echo failure

s1_1    s2_1
s2_1    s1_1
success

# double links also between layers: we expect double links between s2_1 and s3_1
printf ">s1_3\nA\n>s2_1\nC\n>s3_1\nCT\n" | 
swarm -l /dev/null -o /dev/null -j - | \
tail -n 2 | tr "\n" " " | \
awk '{exit $1 == $4 && $2 == $3 ? 0 : 1}' && echo success || echo failure

s1_3    s2_1
s2_1    s3_1
s3_1    s2_1
success

# only with d = 1: we expect no link
printf ">s1_3\nA\n>s2_1\nC\n>s3_1\nG\n" | \
swarm -d 2 -o /dev/null -j - 2> /dev/null && \
echo failure || echo success

Error: A network file can only written when d=1.
success

# using fastidious option: we expect no error
printf ">s1_3\nA\n>s2_1\nCC\n" | swarm -f -l /dev/null -o /dev/null -j - && \
echo success || echo failure

success

# with fastidious option: we expect no link 
# (referred to "test sequences with two differences (different abundances)")
printf ">s1_3\nA\n>s2_1\nCC\n" | swarm -f -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 0 ? 0 : 1}' && echo success || echo failure

success

# fastidious option doesn't change behaviour of -j
printf ">s1_3\nA\n>s2_1\nC\n" | swarm -f -l /dev/null -o /dev/null -j - | \
awk '{exit NR == 1 ? 0 : 1}' && echo success || echo failure

s1_3    s2_1
success

# -n to get full network (abundance values don't matter anymore)
# we expect a double link
printf ">s1_3\nA\n>s2_1\nC\n" | swarm -n -l /dev/null -o /dev/null -j - | tr "\n" " " | \
awk '{exit $1 == $4 && $2 == $3 ? 0 : 1}' && echo success || echo failure

s1_3    s2_1
s2_1    s1_3
success
```

2. -fastidious and -boundary option:
when using the option --fastidious (-f), define the minimum mass of a 
large OTU. By default, an OTU with a mass of 3 or more is considered 
large. Conversely, an OTU is small if it has a mass of less than 3, 
meaning that it is composed of either one amplicon of abundance 2, or 
two amplicons of abundance 1. Any positive value greater than 1 can be 
specified. Using higher boundary values will speed up the second pass, 
but also reduce the taxonomical resolution of swarm results. Default 
mass of a large OTU is 3.
```{r}
#!/bin/bash

# Fastidious option links amplicons with two differences
printf ">s1_3\nAA\n>s2_2\nCC\n" | swarm -f -l /dev/null -i - 

s1_3 s2_2
s1      s2      2       1       1

# Without the fastidious option, the amplicons are not linked
printf ">s1_3\nAA\n>s2_2\nCC\n" | swarm -l /dev/null -i - 

s1_3
s2_2

# cluster total size is 4, largest amplicon has an abundance of 3
printf ">s1_3\nAA\n>s2_1\nAC\n>s3_1\nGG\n" | swarm -f -l /dev/null -i -

s1_3 s2_1 s3_1
s1      s2      1       1       1
s1      s3      2       1       1

# cluster total size is 3, no amplicon is bigger than 2
printf ">s1_2\nAA\n>s2_1\nAC\n>s3_1\nGG\n" | swarm -f -l /dev/null -o -
  
s1_2 s2_1 s3_1

# An OTU is small if it has a mass of less than 3: One Amplicon with abundance 2 - s1 and s2 should be linked
printf ">s1_3\nAA\n>s2_2\nCC\n" | swarm -f -l /dev/null -i - 
  
s1_3 s2_2
s1      s2      2       1       1

# An OTU is small if it has a mass of less than 3: One Amplicon with abundance 2 - s1 and s2 should not be linked
printf ">s1_3\nAA\n>s2_3\nCC\n" | swarm -f -l /dev/null -i - 

s1_3
s2_3

# An OTU is small if it has a mass of less than 3: Two amplicons with abundance 1 - s1, s2 and s3 should be linked
printf ">s1_3\nAA\n>s2_1\nCC\n>s3_1\nGC\n" | swarm -f -l /dev/null -i -

s1_3 s2_1 s3_1
s1      s2      2       1       1
s2      s3      1       1       1

# An OTU is small if it has a mass of less than 3: Three amplicons with abundance 1 - there should be two clusters
printf ">s1_2\nAA\n>s2_1\nCC\n>s3_1\nGC\n>s4_1\nCT\n" | swarm -f -l /dev/null -o -

s2_1 s3_1 s4_1 s1_2

# We can change the default mass value when using -b: s1 and s2 should be linked
printf ">s1_4\nAA\n>s2_3\nCC\n" | swarm -f -b 4 -l /dev/null -o -

s1_4 s2_3

# We can change the default mass value when using -b: s1 and s2 should not be linked
printf ">s1_3\nAA\n>s2_3\nCC\n" | swarm -f -l /dev/null -o -

s1_3
s2_3

# Any positive value greater than 1 can be specified: should cause an error
printf ">s1_4\nAA\n>s2_3\nCC\n" | swarm -f -b 0 -l /dev/null -o -

Error: Illegal boundary specified with -b or --boundary, must be at least 2.


exit 0
```

